Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-08-04 draft.tex
Set utf encoding to spell Marcin's last name
[intro] Start work

@@

\preprint{SI-HEP-2014-17}
\title{Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrzaszcz}
\email{mchrzasz@cern.ch}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

@@

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{^N \theta} f_i(\thvec) f_j(\thvec)  = \delta_{ij}\,,
\end{equation}
with $\Omega$ representing the full angular phase space relevant to the process.
For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays which we consider the most interesting
we list the transformations in a series of appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.\\

@@

In the remainder of this letter we discuss how to obtain the angular
observables $S_i(\nuvec)$ in an experimental setup where each recorded
event is (approximately) distributed according to $P$.  We establish
the statistical basics in section \ref{sec:sample-based-det}. Section
\ref{sec:systematics} is dedicated to the impact of systematic effects
such as mismodeling the underlying physics or detector acceptance
effects. Numerical studies for one uni-angular and one triple-angular
distribution are provided in section \ref{sec:numerics}. In a series
of appendices we provide the orthonormal bases of angular functions
relevant to several rare $b$ decays.

@@

\section{Application to $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ (S-wave and P-wave)}
\label{app:btokstarll}
The PDF for the decay $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ --- up to and including P-wave contributions --- has been calculated
for the most general basis of dimension-six $b\to s$ operators. It reads, expressed in terms of the angular observables $\lbrace J_i\rbrace$ \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_k \frac{J_{k}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_k(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $\theta_1 \equiv \theta_\ell$ is the dilepton helicity angle, and $\theta_2 \equiv \theta_{K}$ is the $\bar{K}\pi$ helicity angle, and $\theta_3 \equiv \phi$ is the azimuthal angle.
The $q^2$-differential decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}}{\rmdx{q^2}} = \frac{\big(3 J_{1c} - J_{2c}\big) + 2\big(3J_{1s} - J_{2s}\big)}{3}\,.
\end{equation}
The angular functions\footnote{%
    We consistently use the label "$i$" to denote angular function and observables that arise
    from interference between S-wave and P-wave amplitudes. Our convention is related to the convention
    of \cite{Bobeth:2012vn} through $f_{1i} \leftrightarrow f_{1sc}$, $f_{2i} \leftrightarrow f_{2sc}$.
} read \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    f_{1s} & = \sin^2\theta_2 &
    f_{1c} & = \cos^2\theta_2 &
    f_{1i} & = \cos\theta_2\\
    f_{2s} & = \sin^2\theta_2 \cos 2\theta_1 &
    f_{2c} & = \cos^2\theta_2 \cos 2\theta_1 &
    f_{2i} & = \cos\theta_2 \cos 2\theta_1\\
    f_{3}  & = \sin^2\theta_2\sin^2\theta_1 \cos 2\theta_3 &
    f_{4}  & = \sin 2\theta_2 \sin 2\theta_1 \cos\theta_3 &
    f_{4i} & = \sin\theta_2 \sin 2\theta_1 \cos\theta_3\\
    f_{5}  & = \sin 2\theta_2 \sin \theta_1 \cos\theta_3 & & &
    f_{5i} & = \sin\theta_2 \sin \theta_1 \cos\theta_3\\
    f_{6s} & = \sin^2\theta_2 \cos\theta_1 &
    f_{6c} & = \cos^2\theta_2 \cos\theta_1\\
    f_{7}  & = \sin 2\theta_2 \sin \theta_1 \sin\theta_3 & & &
    f_{7i} & = \sin\theta_2 \sin \theta_1 \sin\theta_3\\
    f_{8}  & = \sin 2\theta_2 \sin 2\theta_1 \sin\theta_3 & & &
    f_{8i} & = \sin\theta_2 \sin 2\theta_1 \sin\theta_3\\
    f_{9}  & = \sin^2\theta_2\sin^2\theta_1 \sin 2\theta_3\\
\end{aligned}
\end{equation}
We indicate contributions from the interference of S waves and P waves by an index \emph{i}. In terms of the angles $\theta_1$ and $\theta_2$, instead of their cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_k S_k(q^2) f_k(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_k(q^2) \equiv J_k(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

\preprint{SI-HEP-2014-17}
\title{Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays}
\author{Frederik Beaujean}
\email{frederik.beaujean@lmu.de}
\affiliation{C2PAP, Universe Cluster, Ludwig-Maximilians-Universit\"at M\"unchen, Garching, Germany}
\author{Marcin Chrz\k{a}szcz}
\email{mchrzasz@cern.ch}
\author{Nicola Serra}
\email{nicola.serra@cern.ch}
\affiliation{Physik-Institut, Universit\"at Z\"urich, Z\"urich, Switzerland}
\author{Danny van Dyk}
\email{vandyk@tp1.physik.uni-siegen.de}
\affiliation{Theoretische Physik 1, Naturwissenschaftlich-Technische Fakult\"at,
Universit\"at Siegen, Siegen, Germany}

@@

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$\{f_i(\thvec)\}$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{^N \theta} f_i(\thvec) f_j(\thvec)  = \delta_{ij}\,,
\end{equation}
with $\Omega$ representing the full angular phase space relevant to the process.
For the purpose of this letter, however, it suffices that the system
of angular functions $\{f_i(\thvec)\}$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays which we consider the most interesting
we list the transformations in a series of appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.\\

@@

% Our method is closely related to Pearson's method of moments
% \citeneeded{}. One draws samples from a PDF $P(x | \nuvec)$ that
% depends on parameters $\nuvec$. Then one approximates the population
% (central) moments $x^k$ by the sample moments
Our method is an extension of the classical method of moments with
orthogonal functions~\cite{James:2006zz}.

In the remainder of this letter we discuss how to obtain an angular
observable $S_i(\nuvec)$ in an experimental setup where each recorded
event is (approximately) distributed according to $P$.  We establish
the statistical basics in section \ref{sec:sample-based-det}. Section
\ref{sec:systematics} is dedicated to the impact of systematic effects
such as mismodeling the underlying physics or detector acceptance
effects. Numerical studies for one uni-angular and one triple-angular
distribution are provided in section \ref{sec:numerics}. In a series
of appendices we provide the orthonormal bases of angular functions
relevant to several rare $b$ decays.

@@

\section{Application to $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ (S-wave and P-wave)}
\label{app:btokstarll}

The PDF for the decay $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ --- up to and including P-wave contributions --- has been calculated
for the most general basis of dimension-six $b\to s$ operators. It reads, expressed in terms of the angular observables $\lbrace J_i\rbrace$ \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_k \frac{J_{k}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_k(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $\theta_1 \equiv \theta_\ell$ is the dilepton helicity angle, and $\theta_2 \equiv \theta_{K}$ is the $\bar{K}\pi$ helicity angle, and $\theta_3 \equiv \phi$ is the azimuthal angle.
The $q^2$-differential decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}}{\rmdx{q^2}} = \frac{\big(3 J_{1c} - J_{2c}\big) + 2\big(3J_{1s} - J_{2s}\big)}{3}\,.
\end{equation}
The angular functions\footnote{%
    We consistently use the label "$i$" to denote angular function and observables that arise
    from interference between S-wave and P-wave amplitudes. Our convention is related to the convention
    of \cite{Bobeth:2012vn} through $f_{1i} \leftrightarrow f_{1sc}$, $f_{2i} \leftrightarrow f_{2sc}$.
} read \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    f_{1s} & = \sin^2\theta_2 &
    f_{1c} & = \cos^2\theta_2 &
    f_{1i} & = \cos\theta_2\\
    f_{2s} & = \sin^2\theta_2 \cos 2\theta_1 &
    f_{2c} & = \cos^2\theta_2 \cos 2\theta_1 &
    f_{2i} & = \cos\theta_2 \cos 2\theta_1\\
    f_{3}  & = \sin^2\theta_2\sin^2\theta_1 \cos 2\theta_3 &
    f_{4}  & = \sin 2\theta_2 \sin 2\theta_1 \cos\theta_3 &
    f_{4i} & = \sin\theta_2 \sin 2\theta_1 \cos\theta_3\\
    f_{5}  & = \sin 2\theta_2 \sin \theta_1 \cos\theta_3 & & &
    f_{5i} & = \sin\theta_2 \sin \theta_1 \cos\theta_3\\
    f_{6s} & = \sin^2\theta_2 \cos\theta_1 &
    f_{6c} & = \cos^2\theta_2 \cos\theta_1\\
    f_{7}  & = \sin 2\theta_2 \sin \theta_1 \sin\theta_3 & & &
    f_{7i} & = \sin\theta_2 \sin \theta_1 \sin\theta_3\\
    f_{8}  & = \sin 2\theta_2 \sin 2\theta_1 \sin\theta_3 & & &
    f_{8i} & = \sin\theta_2 \sin 2\theta_1 \sin\theta_3\\
    f_{9}  & = \sin^2\theta_2\sin^2\theta_1 \sin 2\theta_3\\
\end{aligned}
\end{equation}
We indicate contributions from the interference of S waves and P waves by an index \emph{i}. In terms of the angles $\theta_1$ and $\theta_2$, instead of their cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_k S_k(q^2) f_k(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_k(q^2) \equiv J_k(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\
